A 38 Million Year Old Mini-Neptune in the Kepler Field

### 2021-09-15 paper/ms.tex
update ms (binary companion appendix)
Co-authored-by: adamkraus <alk@astro.as.utexas.edu>

Beyond the kinematic and rotational lines of evidence for the youth of
\sn, there is a final consistency check: the abundance of
Li\,\textsc{I} as inferred from the 6708\,\AA\ doublet \citep[see {\it
e.g.},][]{soderblom_ages_2014}.  We acquired an iodine-free spectrum
from Keck/HIRES on the night of 26 March 2021 using the standard setup
and reduction techniques of the California Planet Survey
\citep{howard_cps_2010}.  Following the equivalent width measurement
procedure described by \citet{bouma_2021_ngc2516}, we find ${\rm
EW}_{\rm Li} = 233^{+5}_{-7}$\,m\AA.   Note that this value does not
correct for the Fe\,\textsc{I} blend at 6707.44\AA.  Given the
effective temperature of \sn\ (Table~\ref{tab:starparams}), this
measurement is also in line with expectations for a $\approx$38\,Myr
star ({\it e.g.}, \citealt{randich_gaiaeso_2018};
\citealt{bouma_2021_ngc2516}).

@@

		\subfloat{
			\includegraphics[width=0.49\textwidth]{f16.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
	\caption{
		{\bf Evidence for a prograde orbit of Kepler 1627 Ab.}
    {\it Top:} 
    The time of each Kepler transit was measured, along with the local
    slope of the light curve.  The two quantities appear
    anti-correlated, which is most easily explained by starspot
    crossings during the first (second) half of transit inducing a
    positive (negative) TTV, provided that the orbit is prograde
    \citep{mazeh_time_2015}.  The units along the abscissa are most
    easily understood by considering that the stellar flux
    changes by $\sim$60\,ppt per half rotation period
    ($\sim$1.3\,days).
    {\it Bottom:} Radial velocities acquired using Keck/HIRES during
    the transit of 7 Aug 2021 UT, observed simultaneously with MuSCAT3
    ({\it griz} photometry).  A quadratic baseline trend was allowed
    in the RV fit.  The best-fit sky-projected obliquity in this model
    is listed in red; the detection however has marginal statistical
    significance compared to a ``trend-only'' model.
    \label{fig:obliquity}
	}
\end{figure}

@@

The final piece of evidence comes from a spectroscopic transit we
observed using Keck/HIRES during the night of 7 Aug 2021.  We
used the iodine cell for wavelength calibration, and scheduled the
transit using the ephemeris of \citet{holczer_transit_2016}.  We also acquired
simultaneous multi-band photometry using MuSCAT3 \citep{Narita_2020}.
The lower panels in Figure~\ref{fig:obliquity} show the results.  The
photometry shows the expected starspot trend, along with the transit
and a likely starspot crossing event at ${\rm JD} - 2459433 = 0.955$.
The radial velocities show significant scatter, and an
increase in RV is seen shortly after the expected ingress time.  The
internal uncertainty on the RVs ($\approx 15\,{\rm m\,s}^{-1}$) is
dominated by the $v\sin i$ broadening.

@@

\begin{figure}[tp]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.49\textwidth]{f14.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
    {\bf Radii, orbital periods, and ages of known transiting
    exoplanets}.  Planets older than $10^9$ years are shown with gray
    markers.  Planets younger than a gigayear with ${\rm
    Age}/\sigma_{\rm Age} > 3$ are emphasized.  \pn\ is shown with a
    star.  The large radii of the youngest transiting planets could be
    explained by their primordial atmospheres not yet having
    evaporated; direct measurements of the atmospheric outflows or
    planetary masses would help to confirm this expectation.
    Selection effects may also be important.  Parameters are from the
    NASA Exoplanet Archive (July 2, 2021).
		\label{fig:rp_period_age}
	}
\end{figure}

@@

\begin{figure*}[tp]
	\begin{center}
		\leavevmode
		\subfloat{
			\includegraphics[width=0.49\textwidth]{f17a.pdf}
			\includegraphics[width=0.37\textwidth]{f17b.jpg}
		}

		\vspace{-0.1cm}
		\subfloat{
			\includegraphics[width=0.55\textwidth]{f6.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
	\caption{
    {\bf Kepler 1627 is a binary.} {\it Top:} High-resolution imaging
    from Gemini-North/`Alopeke ({\it left}) and Keck/NIRC2 ({\it
    right}) shows an $\approx$M2V companion at $\rho \approx
    0\farcs16$, which corresponds to a projected separation of
    $53\pm4$\,au.  {\it Bottom:} Gaia EDR3 renormalized unit weight
    error (RUWE) point estimates for Kepler\,1627A and other members
    of the $\delta$\,Lyr cluster reported by \citet{KounkelCovey2019}.
    Since other members of the cluster with similar brightnesses have
    comparable degrees of photometric variability, the high RUWE
    suggests that Kepler\,1627 is an astrometric binary. 
    \label{fig:kep1627binary}
	}
\end{figure*}

We first noted the presence of a stellar neighbor in archival
Keck/NIRC2 Kp-band adaptive optics imaging (2.12\,$\mu $m), from the
Spring semesters of both 2015 and 2019.  We acquired additional
Gemini-North/`Alopeke speckle imaging on 24 June 2021, which showed
the same companion in the 0.83\,$\mu$m notch filter, but not the
$0.56\,\mu$m filter.  Figure~\ref{fig:kep1627binary} summarizes the
results of the high-resolution imaging.  Although the companion is
unresolved in the Gaia source catalog, its existence was also
suggested by the primary star's large RUWE, relative to other members
of the \cn.  The images yield a projected separation $\rho \approx
0\farcs16$, with $\Delta m_{832} \approx 3.1$ and $\Delta m_{\rm Kp} =
2.5$.  Using the measured Gaia EDR3 parallax for the system, this
implies a projected separation of $53 \pm 4$\,au.  Given the large
observed RUWE of the primary, and the low geometric probability that a
companion imaged within $\rho \approx 0\farcs16$ is a chance
line-of-sight companion, we proceed under the assumption that the
companion is bound, and that Kepler\,1627 system is a binary.  Based
on the
tabulation\footnote{\url{http://www.pas.rochester.edu/~emamajek/EEM_dwarf_UBVIJHK_colors_Teff.txt},
version \texttt{2021/03/02}.} by \citet{pecaut_mamajek_2013}, the
measured magnitude differences for a G8V primary correspond to a
spectral type for the companion of $\approx$M2V ($M_\star \approx
0.44\,M_\odot$).  However, the companion should have a longer
pre-main-sequence contraction phase than the primary, which would
imply that this mass is overestimated by $\sim10$ to $20\%$.

>>>>

Beyond the kinematic and rotational lines of evidence for the youth of
\sn, there is a final consistency check: the abundance of
Li\,\textsc{I} as inferred from the 6708\,\AA\ doublet \citep[see {\it
e.g.},][]{soderblom_ages_2014}.  We acquired an iodine-free spectrum
from Keck/HIRES on the night of 2021 March 26 using the standard setup
and reduction techniques of the California Planet Survey
\citep{howard_cps_2010}.  Following the equivalent width measurement
procedure described by \citet{bouma_2021_ngc2516}, we find ${\rm
EW}_{\rm Li} = 233^{+5}_{-7}$\,m\AA.   Note that this value does not
correct for the Fe\,\textsc{I} blend at 6707.44\AA.  Given the
effective temperature of \sn\ (Table~\ref{tab:starparams}), this
measurement is also in line with expectations for a $\approx$38\,Myr
star ({\it e.g.}, \citealt{randich_gaiaeso_2018};
\citealt{bouma_2021_ngc2516}).

@@

		\subfloat{
			\includegraphics[width=0.49\textwidth]{f16.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
	\caption{
		{\bf Evidence for a prograde orbit of Kepler 1627 Ab.}
    {\it Top:} 
    The time of each Kepler transit was measured, along with the local
    slope of the light curve.  The two quantities appear
    anti-correlated, which is most easily explained by starspot
    crossings during the first (second) half of transit inducing a
    positive (negative) TTV, provided that the orbit is prograde
    \citep{mazeh_time_2015}.  The units along the abscissa are most
    easily understood by considering that the stellar flux
    changes by $\sim$60\,ppt per half rotation period
    ($\sim$1.3\,days).
    {\it Bottom:} Radial velocities acquired using Keck/HIRES during
    the transit of 2021 Aug 7, observed simultaneously with MuSCAT3
    ({\it griz} photometry).  A quadratic baseline trend was allowed
    in the RV fit.  The best-fit sky-projected obliquity in this model
    is listed in red; the detection however has marginal statistical
    significance compared to a ``trend-only'' model.
    \label{fig:obliquity}
	}
\end{figure}
\begin{figure*}[tp]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.8\textwidth]{f14.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
		{\bf Radii, orbital periods, and ages of transiting
			exoplanets}.  Planets older than $10^9$ years are shown with gray
		markers.  Planets younger than a gigayear with ${\rm
			\tau}/\sigma_{\rm \tau} > 3$ are emphasized, for $\tau$ the age.
		\pn\ is shown with a
		star.  The large sizes of the youngest transiting planets could be
		explained by their primordial atmospheres not yet having
		evaporated; direct measurements of the atmospheric outflows or
		planetary masses would help to confirm this expectation.
		Selection effects may also be important.  Parameters are from the
		NASA Exoplanet Archive (2021 July 2).
		\label{fig:rp_period_age}
	}
\end{figure*}

@@

The final piece of evidence comes from a spectroscopic transit we
observed using Keck/HIRES during the night of 2021 Aug 7.  We
used the iodine cell for wavelength calibration, and scheduled the
transit using the ephemeris of \citet{holczer_transit_2016}.  We also acquired
simultaneous multi-band photometry using MuSCAT3 \citep{Narita_2020}.
The lower panels in Figure~\ref{fig:obliquity} show the results.  The
photometry shows the expected starspot trend, along with the transit
and a likely starspot crossing event at ${\rm JD} - 2459433 = 0.955$.
The radial velocities show significant scatter, and an
increase in RV is seen shortly after the expected ingress time.  The
internal uncertainty on the RVs ($\approx 15\,{\rm m\,s}^{-1}$) is
dominated by the $v\sin i$ broadening.

@@

@@

\begin{figure*}[tp]
	\begin{center}
		\leavevmode
		\subfloat{
			\includegraphics[width=0.49\textwidth]{f17.pdf}
			\includegraphics[width=0.48\textwidth]{f6.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
	\caption{
    {\bf Kepler 1627 is a binary.} {\it Left:} High-resolution imaging
    from Gemini-North/`Alopeke and Keck/NIRC2 shows an $\approx$M2V
    companion at $\rho \approx 0\farcs16$, which corresponds to a
    projected separation of $53\pm4$\,AU.  The inset shows a cutout of
    the NIRC2 image (North is up, East is left, scale is set by the
    separation of the binary).  The lines show 5-$\sigma$ contrast
    limits for the `Alopeke filters, and 6-$\sigma$ contrast limits
    for NIRC2.  {\it Left:} Gaia EDR3 renormalized
    unit weight error (RUWE) point estimates for Kepler\,1627A and
    other members of the $\delta$\,Lyr cluster reported by
    \citet{KounkelCovey2019}.  Since other members of the cluster with
    similar brightnesses have comparable degrees of photometric
    variability, the high RUWE suggests that Kepler\,1627 is an
    astrometric binary. 
    \label{fig:kep1627binary}
	}
\end{figure*}

We first noted the presence of a close neighbor on 2015 July 22 when
we acquired adaptive optics imaging of \sn\ using the NIRC2 imager on
Keck-II. We used the narrow camera (FOV = 10.2\arcsec) to obtain 8
images in the $K'$ filter ($\lambda = 2.12\,\mu$m) with a total
exposure time of 160\,s. We analyzed these data following
\citet{kraus_impact_2016}, beginning by applying standard reduction
steps (dark-subtraction, flat-fielding, linearization, and bad pixel
correction). We then used PSF-fitting to measure the separation,
position angle, and contrast of the candidate companion, identifying
the best-fitting empirical PSF template from among the
near-contemporaneous observations of single stars in the same filter.
Finally, we used the optical distortion solution of \citet{yelda_2010}
to convert pixel-space relative positions to on-sky relative
astrometry. The mean values inferred from the 8 images are reported in
Table~\ref{tab:starparams}.

We also estimated detection limits following
\citet{kraus_impact_2016}, analyzing the residuals after subtracting
the empirical PSF templates, as well as the residuals after
subtracting an azimuthally smoothed PSF profile. Within each residual
image, the flux was measured through 40\,mas apertures centered on
every pixel, and then the noise as a function of radius was estimated
from the RMS within concentric rings. Finally, the detection limits
were estimated from the strehl-weighted sum of the detection
significances in the image stack, and we adopted the $+6\sigma$
threshold as the detection limit for ruling out additional companions.
We find contrast limits of $\Delta m_{K'} = 5.1$ mag at $\rho = 150$ mas,
$\Delta m_{K'} = 7.0$ mag at $\rho = 300$ mas, and $\Delta m_{K'} = 9.0$ mag
at $\rho > 1000$ mas.  Given the distance ($d = 330$\,pc) and age
($\tau \sim 40$\,Myr), the models of \citet{chabrier_2000} imply
corresponding physical limits on tertiary companions of $M_{\rm ter} < 50
M_{\rm Jup}$ at $\rho = 50$\,AU, $M_{\rm ter} < 20 M_{\rm Jup}$ at $\rho =
100$\,AU, and $M_{\rm ter} < 10 M_{\rm Jup}$ at $\rho = 330$\,AU.

We also observed \sn\ on Gemini-North using the `Alopeke speckle
imager on 2021 June 24.  `Alopeke is a dual-channel speckle
interferometer that uses narrow-band filters centered at 0.83\,$\mu$m
and $0.56\,\mu$m.  We acquired three sets of $1000\times 60$$\,$msec
exposures during good seeing (0.45$''$), and used the autocorrelation
function of these images to reconstruct a single image and 5-$\sigma$
detection limits (see \citealt{howell_speckle_2011}).  This procedure
yielded a detection of the companion in the 0.83\,$\mu$m notch filter,
but not the $0.56\,\mu$m filter.
The measured projected separation and magnitude difference
are given in Table~\ref{tab:starparams}.

Figure~\ref{fig:kep1627binary} summarizes the results of the
high-resolution imaging.  
The Gaia EDR3 parallax for the primary implies a projected separation
of $53 \pm 4$\,AU, assuming the companion is bound. 
Although the companion is unresolved in the
Gaia source catalog (there are no candidates brighter than $G < 20.5$ mag within $\rho < 120\arcsec$), its existence was also suggested by the primary
star's large RUWE, relative to other members of the \cn.  Given the
large observed RUWE of the primary, and the low geometric probability
that a companion imaged at $\rho \approx 0\farcs16$ is a chance
line-of-sight companion, we proceed under the assumption that the
companion is bound, and that Kepler\,1627 is a binary. 
Given the distance and age, the models of \citet{baraffe_new_2015} imply
a companion mass of $M_{\rm B} \approx 0.33 M_{\odot}$ and companion
temperature of $T_{\rm eff,B} \approx 3450$ K.
The corresponding spectral type is roughly M2.5V \citep{pecaut_mamajek_2013}.
% \footnote{\url{http://www.pas.rochester.edu/~emamajek/EEM_dwarf_UBVIJHK_colors_Teff.txt}, version \texttt{2021/03/02}.}
